---
title: The title
output:
  bookdown::html_document2:
    toc: yes
    toc_depth: 2
    number_sections: yes
    fig_width: 5
    fig_height: 5
    highlight: zenburn
    df_print: kable
    code_folding: hide
    theme: spacelab
  bookdown::word_document2:
    toc: yes
    toc_depth: 2
    number_sections: yes
    fig_width: 5
    fig_height: 5
    highlight: zenburn
    df_print: kable	
  bookdown::pdf_document2:
    latex_engine: xelatex
    toc: yes
    toc_depth: 2
    number_sections: yes
    fig_width: 5
    fig_height: 5
    highlight: zenburn
    df_print: kable
documentclass: article
mainfont: Arial
...

# Section 1
Text with embedded R code.

```{r Summary}
mean(rnorm(10))
```
```{r head}
head(cars)
```

## Subsection 1						

We can include a figure

```{r Plot, fig.cap='Gaussian density.',message=FALSE}
library(tidyverse)
data.frame(x=rnorm(1000)) %>%
	ggplot(aes(x=x)) + geom_density()
```

Perhaps even a table

```{r table}
library(knitr)
kable(summary(cars), caption='Summary of the cars data set')
```
