---
title: "Data wrangling"
author: "Murray Logan"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    code_folding: show
    collapse: no
    df_print: paged
    fig_caption: yes
    fig_height: 4
    fig_width: 4
    highlight: textmate
    theme: spacelab
    toc: yes
    toc_float: yes
    css: ../resources/style.css
  pdf_document:
    df_print: default
    fig_caption: yes
    fig_height: 4
    fig_width: 4
    highlight: tango
    latex_engine: xelatex
    number_sections: yes
    toc_depth: 2
  word_document:
    fig_caption: yes
    fig_height: 4
    fig_width: 4
    highlight: tango
    toc: yes
    toc_depth: 2
output_dir: "docs"
documentclass: article
fontsize: 12pt
mainfont: Arial
mathfont: LiberationMono
monofont: DejaVu Sans Mono
classoption: a4paper
bibliography: ../resources/references.bib
---

```{r setup, include=FALSE, warnings=FALSE, message=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
 
# Preparations

Load the necessary libraries

```{r libraries, results='markdown', eval=TRUE, message=FALSE, warning=FALSE}
library(tidyverse) #for data wrangling
```

Link to the data transformation cheatsheet 

https://github.com/rstudio/cheatsheets/raw/master/data-transformation.pdf

Important data manipulation functions:

+--------------------------+-----------------------------------+------------+
|Task                      |Function                           |Package     |
+==========================+===================================+============+
|Sorting                   |`arrange()`                        |**dplyr**   |
+--------------------------+-----------------------------------+------------+
|Adding columns            |`mutate()`                         |**dplyr**   |
+--------------------------+-----------------------------------+------------+
|Transformations           |`mutate()`                         |**dplyr**   |
+--------------------------+-----------------------------------+------------+
|Re-ordering factor levels |`factor(,levels=)`                 |base        |
+--------------------------+-----------------------------------+------------+
|Re-labelling              |`factor(,lab=)`                    |base        |
+--------------------------+-----------------------------------+------------+
|                          |`recode()`                         |**dplyr**   |
+--------------------------+-----------------------------------+------------+
|Re-naming columns         |`rename(,replace=)`                |**dplyr**   |
+--------------------------+-----------------------------------+------------+
|Filtering/Subsetting      |indexing                           |base        |
+--------------------------+-----------------------------------+------------+
|~ columns                 |`select(,...)`                     |**dplyr**   |
+--------------------------+-----------------------------------+------------+
|                          |`pull(,...)`                       |**dplyr**   |
+--------------------------+-----------------------------------+------------+
|~ rows                    |`filter(,...)`                     |**dplyr**   |
+--------------------------+-----------------------------------+------------+
|Unique combinations       |`distinct()`                       |**dplyr**   |
+--------------------------+-----------------------------------+------------+
|Reshaping data            | `pivot_longer()`, `pivot_wider()` | **tidyr**  |
+--------------------------+-----------------------------------+------------+
|Split/combine columns     | `separate()`, `unite()`           | **tidyr**  |
+--------------------------+-----------------------------------+------------+
|Aggregating               |`group_by()` `summarise()`         |**dplyr**   |
+--------------------------+-----------------------------------+------------+
|                          |`group_by()` `count()`             |**dplyr**   |
+--------------------------+-----------------------------------+------------+
|Merging/joining           |`*_join()`                         |**dplyr**   |
+--------------------------+-----------------------------------+------------+
|Extracting data structure |`expand()`                         |**tidyr**   |
+--------------------------+-----------------------------------+------------+
|                          |`crossing()`                       |**tidyr**   |
+--------------------------+-----------------------------------+------------+


# Piping

# Data files
 
```{r getData, results='markdown', eval=TRUE}
load(file='../data/manipulationDatasets.RData')
dat.1 %>% head
```

Sorting data
================
```{r, results='markup'}
head(dat.1)
#OR
dat.1 %>% head()
#OR
dat.1 %>% head
```

Adding columns - mutate
===========================


Summarising (aggregating) data
=================================


Grouping (=aggregating)
=========================


Subset columns
=================

## Regular expressions (regex)

https://github.com/rstudio/cheatsheets/raw/master/data-transformation.pdf

Filtering
=============

Reshaping data
=================

## Pivot longer

## Pivot wider

Combining data
=================

Applied examples
===================
