---
title: "Data wrangling"
author: "Murray Logan"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    code_folding: show
    collapse: no
    df_print: paged
    fig_caption: yes
    fig_height: 4
    fig_width: 4
    highlight: textmate
    theme: spacelab
    toc: yes
    toc_float: yes
    css: ../resources/style.css
  pdf_document:
    df_print: default
    fig_caption: yes
    fig_height: 4
    fig_width: 4
    highlight: tango
    latex_engine: xelatex
    number_sections: yes
    toc_depth: 2
  word_document:
    fig_caption: yes
    fig_height: 4
    fig_width: 4
    highlight: tango
    toc: yes
    toc_depth: 2
output_dir: "docs"
documentclass: article
fontsize: 12pt
mainfont: Arial
mathfont: LiberationMono
monofont: DejaVu Sans Mono
classoption: a4paper
bibliography: ../resources/references.bib
---

```{r setup, include=FALSE, warnings=FALSE, message=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Preparations

Load the necessary libraries

```{r libraries, results='markdown', eval=TRUE, message=FALSE, warning=FALSE}
library(tidyverse) #for data wrangling
```

Link to the data transformation cheatsheet

https://github.com/rstudio/cheatsheets/raw/master/data-transformation.pdf

Important data manipulation functions:

| Task                      | Function                    | Package   |
|---------------------------|-----------------------------|-----------|
| Sorting                   | `arrange()`                 | **dplyr** |
| Re-ordering factor levels | `factor()`                  | **base**  |
| Re-labelling              | `recode()`                  | **dplyr** |
| Re-naming columns         | `rename()`                  | **dplyr** |
| Filtering/subsetting      | `select()`                  | **dplyr** |
| Transformations           | `mutate()`                  | **dplyr** |
| Adding columns            | `mutate()`                  | **dplyr** |
| Re-shaping data           | `gather()`/`spread()`       | **tidyr** |
| Aggregating               | `summarize()`, `group_by()` | **dplyr** |
| Merging/joining           | `*_join()`                  | **dplyr** |
|                           |                             |           |

# Piping

# Data files

```{r getData, results='markdown', eval=TRUE}
load(file='../data/manipulationDatasets.RData')
data.1 %>% head
```

Sorting data
================


Manipulating factors
=======================


Subset columns
=================

## Regular expressions (regex)

https://github.com/rstudio/cheatsheets/raw/master/data-transformation.pdf


Filtering
=============

Adding columns - mutate
===========================


Summarising (aggregating) data
=================================

Piping
=======

Grouping (=aggregating)
=========================

Reshaping data
=================

## Gathering

## Spread

Combining data
=================

Applied examples
===================
