---
title: "GLM example 9"
author: "Murray Logan"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    code_folding: show
    collapse: no
    df_print: paged
    fig_caption: yes
    fig_height: 4
    fig_width: 4
    highlight: textmate
    theme: spacelab
    toc: yes
    toc_float: yes
    css: ../resources/style.css
  pdf_document:
    df_print: default
    fig_caption: yes
    fig_height: 4
    fig_width: 4
    highlight: tango
    latex_engine: xelatex
    number_sections: yes
    toc_depth: 2
  word_document:
    fig_caption: yes
    fig_height: 4
    fig_width: 4
    highlight: tango
    toc: yes
    toc_depth: 2
output_dir: "docs"
documentclass: article
fontsize: 12pt
mainfont: Arial
mathfont: LiberationMono
monofont: DejaVu Sans Mono
classoption: a4paper
bibliography: ../resources/references.bib
---

```{r setup, include=FALSE, warnings=FALSE, message=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Preparations

Load the necessary libraries

```{r libraries, results='markdown', eval=TRUE, message=FALSE, warning=FALSE}
library(car)       #for regression diagnostics
library(broom)     #for tidy output
library(ggfortify) #for model diagnostics
library(sjPlot)    #for outputs
library(knitr)     #for kable
library(effects)   #for partial effects plots
library(DHARMa)    #for residual diagnostics
library(emmeans)   #for estimating marginal means
library(ggeffects) #for diagnostic plots in ggplotjk pR
library(MASS)      #for glm.nb
library(MuMIn)     #for AICc
library(pscl)      #for zero inflated models
library(glmmTMB)
library(nlme)
library(tidyverse) #for data wrangling
library(modelr)    #for auxillary modelling functions
```

# Scenario

To investigate visitor fishing habits in a state park, some wildlife biologists
surveyed the visitors and asked them how many fish they had caught, how many
people were in their group, the number of children in their group and whether
they had brought a camper (caravan) to the park.  As a possible oversight, they
did not ask the visitors whether they had attempted to go fishing and they were
rightly concerned that this might compilate their investigation.

Format of fish.csv data files

CAMPER  CHILDREN  PERSONS    COUNT
------- --------- ---------- ------
0       0         1          0
1       0         1          0
0       0         1          0
1       1         2          0
0       0         1          1

------------------ --------------------------------------------------------------------------------------------
**CAMPER**         Categorical listing of whether or not the visitor had brought
                   a camper to the park ­ independent variable
**CHILDREN**       Categorical listing of the number of children in the
                   visitor's group ­ independent variable
**PERSONS**        The number of people in the visitor's group ­ response variable
**COUNT**          The number of fish the visitor had caught
------------------ --------------------------------------------------------------------------------------------


# Read in the data

```{r readData, results='markdown', eval=TRUE}
fish = read_csv('../data/fish1.csv', trim_ws=TRUE)
glimpse(fish)
```
# Exploratory data analysis


# Fit the model



# Model validation

 



# Model investigation / hypothesis testing




# Predictions



fish.glm1 <- glmmTMB(COUNT ~ CHILDREN * CAMPER, ziform=~PERSONS, data=fish,
                family=poisson(link='log'))

fish.resids <- simulateResiduals(fish.glm1, integerResponse = TRUE)
plot(fish.resids)
testZeroInflation(fish.resids)
testDispersion(fish.resids)

#library(pscl)
#fish.glm2 <- zeroinfl(COUNT ~ CHILDREN * CAMPER | PERSONS,
#                      data=fish,  dist='negbin')

#fish.glm2 <- zeroinfl(COUNT ~ CHILDREN * CAMPER | 1,
#                      data=fish,  dist='negbin')
#summary(fish.glm2)

#fish.glm2 <- glmmTMB(COUNT ~ CHILDREN + CAMPER, ziform=~1, data=fish %>% filter(CHILDREN!='3') %>% droplevels,
#                     family=nbinom2(link='log'))
fish.glm2 <- glmmTMB(COUNT ~ CHILDREN * CAMPER, ziform=~PERSONS, data=fish,
                     family=nbinom2(link='log'))
fish.resids <- simulateResiduals(fish.glm2, integerResponse = TRUE)
plot(fish.resids)
ggemmeans(fish.glm2,  ~CHILDREN|CAMPER) %>% plot

ggemmeans(fish.glm2,  ~CHILDREN|CAMPER) %>% plot

summary(fish.glm2)


fish.glm3 <- glmmTMB(COUNT ~ CHILDREN + CAMPER, ziform=~PERSONS, data=fish,
                     family=nbinom2(link='log'))
#testZeroInflation(fish.resids)
fish.resids <- simulateResiduals(fish.glm3, integerResponse = TRUE)
plot(fish.resids)
AICc(fish.glm2, fish.glm3)

ggemmeans(fish.glm3,  ~CHILDREN|CAMPER) %>% plot

summary(fish.glm3)

emmeans(fish.glm3, pairwise~CHILDREN, type='response') %>%
  confint

newdata <- emmeans(fish.glm3, ~CHILDREN|CAMPER, type='response') %>%
  as.data.frame
ggplot(newdata,  aes(y=response,  x=CHILDREN)) +
  geom_pointrange(aes(ymin=lower.CL,  ymax=upper.CL,  color=CAMPER)) +
  theme_classic()
```
