---
title: "GLMM example 5"
author: "Murray Logan"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    code_folding: show
    collapse: no
    df_print: paged
    fig_caption: yes
    fig_height: 4
    fig_width: 4
    highlight: textmate
    theme: spacelab
    toc: yes
    toc_float: yes
    css: ../resources/style.css
  pdf_document:
    df_print: default
    fig_caption: yes
    fig_height: 4
    fig_width: 4
    highlight: tango
    latex_engine: xelatex
    number_sections: yes
    toc_depth: 2
  word_document:
    fig_caption: yes
    fig_height: 4
    fig_width: 4
    highlight: tango
    toc: yes
    toc_depth: 2
output_dir: "docs"
documentclass: article
fontsize: 12pt
mainfont: Arial
mathfont: LiberationMono
monofont: DejaVu Sans Mono
classoption: a4paper
bibliography: ../resources/references.bib
---

```{r setup, include=FALSE, warnings=FALSE, message=FALSE}
knitr::opts_chunk$set(echo = TRUE, message=FALSE, warning=FALSE,cache.lazy = FALSE, tidy='styler')
```

# Preparations
 
Load the necessary libraries

```{r libraries, results='markdown', eval=TRUE, message=FALSE, warning=FALSE}
library(car)       #for regression diagnostics
library(broom)     #for tidy output
library(broom.mixed) #for tidy output
library(ggfortify) #for model diagnostics
library(sjPlot)    #for outputs
library(knitr)     #for kable
library(effects)   #for partial effects plots
library(ggeffects) #for effects plots in ggplotjk
library(emmeans)   #for estimating marginal means
library(MASS)      #for glm.nb
library(MuMIn)     #for AICc
library(tidyverse) #for data wrangling
library(DHARMa)    #for assessing dispersion etc
library(glmmTMB)    #for glmmTMB
library(performance) #for diagnostic plots
library(see)         #for diagnostic plots
library(lme4)       #for glmer
library(glmmTMB)    #for glmmTMB
```

# Scenario

Some ornithologists were interested in the degree of sibling negotiations in owl
chicks.  Specifically, they wanted to explore how sibling negotiations were
affected by feeding satiety and the sex of the parent returning to the nest.
The ornithologists had accessed to a number of owl nests and were able to count
(via recording equipment) the number of sibling negotiations (calls) that the
owl chicks made when the parent returned to the nest.

We could hypothesise that the chicks might call more if they were hungry.  As
part of the investigation, the researchers were able to provided supplementary
food.  As such, they were able to manipulate the conditions such that sometimes
the chicks in a nest would be considered deprived of supplementary food and at
other times they were satiated.  

As a parent returned, the researchers recorded the number of sibling
negotiations (calls) along with the sex of the parent.  Since the number of
calls is likely to be a function of the number of chicks (the more chicks the
more calls), the researchers also counted the number of siblings in the brood. 

Each nest was measured on multiple occasions.  Hence, we must include the nest
as a random effect to account for the lack of independence between observations
on the same set of siblings.

# Read in the data

```{r readData, results='markdown', eval=TRUE}
owls = read_csv('../data/owls.csv', trim_ws=TRUE)
glimpse(owls)
```

# Data preparation

# Exploratory data analysis


Model formula:
$$
y_i \sim{} \mathcal{Pois}(\lambda_i)\\
ln(\lambda_i) =\boldsymbol{\beta} \bf{X_i} + \boldsymbol{\gamma} \bf{Z_i}
$$

where $\boldsymbol{\beta}$ and $\boldsymbol{\gamma}$ are vectors of the fixed
and random effects parameters respectively and $\bf{X}$ is the model matrix
representing the overall intercept and effects of food treatment, sex of parent,
arrival time (and various interactions) on the number of sibling negotiations.
Brood size was also incorporated as an offset.  $\bf{Z}$ represents a cell means
model matrix for the random intercepts associated with individual nests.


# Fit the model {.tabset .tabset-faded}


# Model validation {.tabset .tabset-faded}



# {-}

# Model refit and validation {.tabset .tabset-faded}



# Partial plots {.tabset .tabset-faded}


# Model investigation / hypothesis testing {.tabset .tabset-faded}


# Predictions


# Summary figures


# References

