---
title: "GLMM example 5"
author: "Murray Logan"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    code_folding: show
    collapse: no
    df_print: paged
    fig_caption: yes
    fig_height: 4
    fig_width: 4
    highlight: textmate
    theme: spacelab
    toc: yes
    toc_float: yes
    css: ../resources/style.css
  pdf_document:
    df_print: default
    fig_caption: yes
    fig_height: 4
    fig_width: 4
    highlight: tango
    latex_engine: xelatex
    number_sections: yes
    toc_depth: 2
  word_document:
    fig_caption: yes
    fig_height: 4
    fig_width: 4
    highlight: tango
    toc: yes
    toc_depth: 2
output_dir: "docs"
documentclass: article
fontsize: 12pt
mainfont: Arial
mathfont: LiberationMono
monofont: DejaVu Sans Mono
classoption: a4paper
bibliography: ../resources/references.bib
---

```{r setup, include=FALSE, warnings=FALSE, message=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Preparations
 
Load the necessary libraries

```{r libraries, results='markdown', eval=TRUE, message=FALSE, warning=FALSE}
library(car)       #for regression diagnostics
library(broom)     #for tidy output
library(broom.mixed) #for tidy output
library(ggfortify) #for model diagnostics
library(sjPlot)    #for outputs
library(knitr)     #for kable
library(effects)   #for partial effects plots
library(ggeffects) #for effects plots in ggplotjk
library(emmeans)   #for estimating marginal means
library(MASS)      #for glm.nb
library(MuMIn)     #for AICc
library(tidyverse) #for data wrangling
library(DHARMa)    #for assessing dispersion etc
library(glmmTMB)    #for glmmTMB
library(performance) #for diagnostic plots
library(see)         #for diagnostic plots
library(lme4)       #for glmer
library(glmmTMB)    #for glmmTMB
```

# Scenario

Some ornithologists were interested in the degree of sibling negotiations in owl
chicks.  Specifically, they wanted to explore how sibling negotiations were
affected by feeding satiety and the sex of the parent returning to the nest.
The ornithologists had accessed to a number of owl nests and were able to count
(via recording equipment) the number of sibling negotiations (calls) that the
owl chicks made when the parent returned to the nest.

We could hypothesise that the chicks might call more if they were hungry.  As
part of the investigation, the researchers were able to provided supplimentary
food.  As such, they were able to manipulate the conditions such that sometimes
the chicks in a nest would be considered deprived of supplimentary food and at
other times they were satiated.  

As a parent returned, the researchers recorded the number of sibling
negotiations (calls) along with the sex of the parent.  Since the number of
calls is likely to be a function of the number of chicks (the more chicks the
more calls), the researchers also counted the number of siblings in the brood. 

Each nest was measured on multiple occasions.  Hence, we must include the nest
as a random effect to account for the lack of independence between observations
on the same set of siblings.

# Read in the data

```{r readData, results='markdown', eval=TRUE}
owls = read_csv('../data/owls.csv', trim_ws=TRUE)
glimpse(owls)
```

# Data preparation

# Exploratory data analysis


Model formula:
$$
y_i \sim{} \mathcal{Pois}(\lambda_i)\\
ln(\lambda_i) =\boldsymbol{\beta} \bf{X_i} + \boldsymbol{\gamma} \bf{Z_i}
$$

where $\boldsymbol{\beta}$ and $\boldsymbol{\gamma}$ are vectors of the fixed and random effects parameters respectively 
and $\bf{X}$ is the model matrix representing the overall intercept and effects of food treatment, sex of parent, arrival time (and various interactions) on the number of sibling negotiations.
Brood size was also incorporated as an offset.
$\bf{Z}$ represents a cell means model matrix for the random intercepts associated with individual nests.


# Fit the model {.tabset .tabset-faded}


# Model validation {.tabset .tabset-faded}



# {-}
<div class='HIDEN'>

**Conclusions:**

- there is evidence that the model does not fit that well. It is evidently zero
  inflated and possibly also overdispersed.
- it would seem that a zero-inflated Poisson or even a zero-inflated Negative
  Binomial would be a sensible next step.
- zero-inflated models cannot be fit in `glmer()`, so we will proceed with
`glmmTMB()` only.

</div>

# Model refit and validation {.tabset .tabset-faded}



# Partial plots {.tabset .tabset-faded}

## plot_model

```{r partialPlots1a, results='markdown', eval=TRUE}
plot_model(owls.glmmTMB4,  type='eff',  terms=c('FoodTreatment', 'SexParent'))
plot_model(owls.glmmTMB6,  type='eff',  terms=c('FoodTreatment', 'SexParent'))
```

These predictions appear to be based on the mean BroodSize of approximately
`r round(mean(owls$BroodSize), 2)`.

## allEffects

```{r partialPlots1b, results='markdown', eval=TRUE}
plot(allEffects(owls.glmmTMB4), multiline=TRUE, ci.style='bars')
```

These predictions also appear to be based on the mean BroodSize, although the
documentation seems to suggest that `allEffects()` might not deal with the
offsets the way we have used them (as a function in the formula) correctly.

## ggpredict

```{r partialPlots1c, results='markdown', eval=TRUE}
ggpredict(owls.glmmTMB4,  terms=c('FoodTreatment', 'SexParent')) %>% plot
```

**This seems to deal with the offset incorrectly**.  For the purpose of
prediction, the offset seems to be set at the value of the first BroodSize (on
the response scale).  This is incorrect for two reasons:

1. it should be on the log scale
2. it would be better to use either the mean BroodSize (then on the link scale)
   or a value of 0 (so as to reflect the per unit BroodSize prediction).
   
## ggemmeans

`ggemmeans()` can accommodate the offset correctly.  There are two sensible
choices:

- set the offset to the (log) of the mean BroodSize (similar to other partial
  effects), hence giving predictions appropriate for the average brood size conclusion.
  
```{r partialPlots1d1, results='markdown', eval=TRUE}
#off<-owls %>% group_by(SexParent, FoodTreatment) %>% summarize(Mean=mean(BroodSize))
off<-owls %>% summarize(Mean=mean(BroodSize))
as.numeric(off)
ggemmeans(owls.glmmTMB4,  ~FoodTreatment+SexParent, offset=log(off$Mean)) %>% plot
```

- set the offset to 0.  This results in predictions appropriate for a per owl
  chick conclusion.

```{r partialPlots1d2, results='markdown', eval=TRUE}
ggemmeans(owls.glmmTMB4,  ~FoodTreatment+SexParent, offset=0) %>% plot
ggemmeans(owls.glmmTMB6,  ~FoodTreatment+SexParent, offset=0) %>% plot
```


# Model investigation / hypothesis testing {.tabset .tabset-faded}


# Predictions


# Summary figures

```{r summaryFig1a, results='markdown', eval=TRUE, hidden=TRUE}
## owls.grid = with(owls, list(FoodTreatment=levels(FoodTreatment),
##                             SexParent=levels(SexParent)))
## newdata = emmeans(owls.glmmTMB4, ~FoodTreatment+SexParent, at=owls.grid,
##                   offset=0, type='response') %>%
##     as.data.frame
newdata = emmeans(owls.glmmTMB4, ~FoodTreatment+SexParent, 
                  offset=0, type='response') %>%
    as.data.frame
head(newdata)
ggplot(newdata, aes(y=rate, x=FoodTreatment)) +
  geom_pointrange(aes(ymin=lower.CL, ymax=upper.CL, color=SexParent),
                  position=position_dodge(width=0.2)) +
  scale_y_continuous('Number of sibling negotiations per chick') +
  theme_bw()

##OR if we want to express this for the average brood size
## owls.grid = with(owls, list(FoodTreatment=levels(FoodTreatment),
##                             SexParent=levels(SexParent)))
## newdata = emmeans(owls.glmmTMB4, ~FoodTreatment+SexParent, at=owls.grid,
##                   offset=log(mean(owls$BroodSize)), type='response') %>%
##     as.data.frame
newdata = emmeans(owls.glmmTMB4, ~FoodTreatment+SexParent,
                  offset=log(mean(owls$BroodSize)), type='response') %>%
    as.data.frame
head(newdata)
ggplot(newdata, aes(y=rate, x=FoodTreatment)) +
  geom_pointrange(aes(ymin=lower.CL, ymax=upper.CL, color=SexParent),
                  position=position_dodge(width=0.2)) +
  scale_y_continuous('Number of sibling negotiations per nest') +
  theme_bw()
```


```{r comparemodels, results='markdown', eval=TRUE, hidden=TRUE, fig.width=10, fig.height=6}

newdata=tidy(owls.glmmTMB3, effects='fixed', conf.int=TRUE,  exponentiate=TRUE) %>%
  mutate(Model='zip (simple zi)') %>%
  bind_rows(
    tidy(owls.glmmTMB4, effects='fixed', conf.int=TRUE,  exponentiate=TRUE) %>%
    mutate(Model='zip (complex zi)')
  ) %>%
  bind_rows(
    tidy(owls.glmmTMB5, effects='fixed', conf.int=TRUE,  exponentiate=TRUE) %>%
    mutate(Model='zinb (simple zi)')
  ) %>%
  bind_rows(
    tidy(owls.glmmTMB6, effects='fixed', conf.int=TRUE,  exponentiate=TRUE) %>%
    mutate(Model='zinb (complex zi)')
  ) %>%
  mutate(Model=factor(Model,  levels=c('zip (simple zi)', 'zip (complex zi)',
                                       'zinb (simple zi)', 'zinb (complex zi)')),
         Cond=interaction(component, term)) %>%
  arrange(component, term) %>%
  mutate(Cond=factor(Cond,  levels=rev(unique(Cond))))

ggplot(newdata,  aes(y=estimate,  x=Cond,  color=Model)) +
  geom_pointrange(aes(ymin=conf.low,  ymax=conf.high),  position=position_dodge(width=0.2)) +
  coord_flip()


newdata = emmeans(owls.glmmTMB3, ~FoodTreatment+SexParent, offset=0, type='response') %>%
  as.data.frame %>% mutate(Model='zip (simple zi)',  response=rate) %>%
  bind_rows(
    emmeans(owls.glmmTMB4, ~FoodTreatment+SexParent, offset=0, type='response') %>%
    as.data.frame %>% mutate(Model='zip (complex zi)',  response=rate)
  ) %>%
  bind_rows(
    emmeans(owls.glmmTMB5, ~FoodTreatment+SexParent, offset=0, type='response') %>%
    as.data.frame %>% mutate(Model='zinb (simple zi)',  response=response)
  ) %>%
  bind_rows(
    emmeans(owls.glmmTMB6, ~FoodTreatment+SexParent, offset=0, type='response') %>%
    as.data.frame %>% mutate(Model='zinb (complex zi)',  response=response)
  ) %>%
  mutate(Model=factor(Model,  levels=c('zip (simple zi)', 'zip (complex zi)',
                                       'zinb (simple zi)', 'zinb (complex zi)')))

head(newdata)
ggplot(newdata, aes(y=response, x=FoodTreatment)) +
  geom_pointrange(aes(color=SexParent, ymin=lower.CL, ymax=upper.CL), 
                  position=position_dodge(width=0.2)) +
  facet_wrap(~Model,  nrow=1)

ggplot(newdata,  aes(y=response,  x=interaction(FoodTreatment,SexParent),  color=Model)) +
  geom_pointrange(aes(ymin=lower.CL,  ymax=upper.CL),  position=position_dodge(width=0.2)) +
  coord_flip()
```

# References

