---
title: "Blank Template"
author: "Murray Logan"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    code_folding: show
    collapse: no
    df_print: paged
    fig_caption: yes
    fig_height: 4
    fig_width: 4
    highlight: textmate
    theme: spacelab
    toc: yes
    toc_float: yes
    css: ../public/resources/style.css
  pdf_document:
    df_print: default
    fig_caption: yes
    fig_height: 4
    fig_width: 4
    highlight: tango
    latex_engine: xelatex
    number_sections: yes
    toc_depth: 2
  word_document:
    fig_caption: yes
    fig_height: 4
    fig_width: 4
    highlight: tango
    toc: yes
    toc_depth: 2
output_dir: "docs"
documentclass: article
fontsize: 12pt
mainfont: Arial
mathfont: LiberationMono
monofont: DejaVu Sans Mono
classoption: a4paper
bibliography: ../public/resources/references.bib
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(tinytex.engine = 'xelatex')
```
   
```{r readRScripts, results='markdown', eval=TRUE, echo=FALSE}
knitr::read_chunk('01_analysis.R')
```   
   
# Preparations

Load the necessary libraries

```{r loadLibraries, eval=TRUE, warning=TRUE, message=FALSE}
```

# Algorithmic approaches

## R-mode analyses

We will use a single common dataset across a whole range of analyses.  Whilst this does
mean that we do not need to keep introducing new datasets, it does give the impression
that each of the analyses are equally suitable for this same dataset.  That is not necessarily
the case...

```{r ReadData}
```

### Principle Components Analysis

#### Unconstrained

#### Constrained (RDA)

### Correspondance Analysis

#### Unconstrained

#### Constrained

### Principle Coordinates Analysis

#### Unconstrained

#### Constrained

## Q-mode analyses

### Multivariate analysis (NMDS)

# Model based approach

# Production information

```{r}
sessionInfo()
```

